Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2024-12-06 draft.tex
minor updates

>>>>

    Lastly, we show a modification of this prove when $p=2$. 
    We choose $\mathring w \in W^{2}\Lambda^{k}(\widetilde \Makoto)$ to have minimum norm among those members of $W^{2}\Lambda^{k}(\widetilde \Makoto)$ whose exterior derivative equals $\cartan \tilde u$. Since then $\tilde w \in W^{2}\Lambda^{k}(\widetilde \Makoto)$ has the same properties,
    it follows that $\tilde w = \mathring w$.

    Under these conditions, we can use Gaffney's inequality: % TODO: cite specific inequality 
    $\tilde w \in W^{1,2}(\Omega) \otimes \Alt^{k}(\bbR^{n})$ with 
    \begin{align}\color{red}
        \| \nabla \tilde w \|_{L^{2}(\Delta^{n})}
        \leq 
        \| \cartan \tilde w \|_{L^{2}(\Delta^{n})}
        .
    \end{align} % TODO: what exactly is the norm being measured on the left hand side?

### 2024-12-06 main.tex
replace some tabs
minor updates

\begin{proof}
    The diameter of $F$ is at least as large as the height $h_S$ of some other vertex of $F$ in $T_{1}$. 
    Now, 
    \begin{align*}
%        \diam(T) \frac{ n! \vol(T) }{ \diam(T)^{n} } \leq h_S \leq \diam(F).
        \diam(T_{1}) \aspectratio(T_{1})^{-1} \leq \height(T_{1}) \leq h_S \leq \diam(F).
    \end{align*}
    The first estimate follows. 
    As for the second estimate,
    let $h_1$ and $h_2$ be the heights of $F$ in the simplices $T_1$ and $T_2$, respectively. 
    By the volume formula for simplices, $\vol(T_1) = h_1 \vol(F)/n$ and $\vol(T_2) = h_2 \vol(F)/n$, and thus $\vol(T_1)/\vol(T_2) = h_1/h_2$. Thus follows the second estimate: 
    \begin{align*}
        \frac{ \vol(T_1) }{ \vol(T_2) }
        = 
        \frac{ h_1 }{ h_2 }
        \leq
        \frac{ \diam(T_1) }{ h_2 }
        \leq
        \aspectratio(T_{1}) \frac{ \diam(F) }{ h_2 }
        \leq
        \aspectratio(T_{1}) \aspectratio(T_{2})
        .
    \end{align*}
    Lastly, let $\Xi : T_1 \rightarrow T_2$ be as stated. 
    We estimate the Lipschitz constant of $\Xi$ as follows. 
    Without loss of generality, $F$ lies in the span of the first $n-1$ coordinates and contains the origin, 
	and the unit height vector $h_0$ goes in the direction of the vertex $z_1$ is in the $n$-th coordinate direction.
    Suppose that $b, b' \in F$ and $\alpha, \alpha' \in \bbR$ such that $x = b + \alpha h_0$ and $x' = b' + \alpha' h_0$.
    Since $\Xi(h_0) = z_2$, it follows with the arguments as above and H\"older's inequality that 
    \begin{align*}
        \| \Xi(x) - \Xi(x') \|
        \leq  
        \| b - b' \| + | \alpha - \alpha' | \cdot \| z_2 \|
        \leq  
        \| b - b' \| + | \alpha - \alpha' | \aspectratio(T_1) \frac{\diam(T_2)}{\diam(T_1)} \| h_0 \|
        \leq 
        \left( 1 + \aspectratio(T_1)^2 \frac{\diam(T_2)^2}{\diam(T_1)^2} \right)^{\frac 1 2} \| x-x' \|
        .
    \end{align*}
    % Writing $h_0$ for the unit normal in direction $z_1$, we have 
    % \begin{align*}
        % \Xi(x) 
        % = 
        % x 
        % - \frac{ \langle h_0, x \rangle }{ \langle h_0, z_1 \rangle } z_1
        % + \frac{ \langle h_0, x \rangle }{ \langle h_0, z_1 \rangle } z_2
        % ,
        % \qquad 
        % x \in \bbR^{n}
        % .
    % \end{align*}
    % Its Jacobian is the rank-one perturbation of the linear projection along $z_1$ onto the hyperplane spanned by $F$. 
    % By the Xu-Zikanatov estimate,
    % \begin{align*}
        % \matnorm{ \Jacobian \Xi }
        % \leq 
        % \frac{ \diam(T_1) }{ \height(T_1) }
        % +
        % \frac{ \diam(T_2) }{ \height(T_1) }
        % \leq 
        % \left( 1 + \aspectratio(T_2) \right)
        % \aspectratio(T_1)
        % .
    % \end{align*}
    The desired estimates are shown.
\begin{comment}
    \mwl{Tighter estimate via geometry}
    Matrix inverse lemma:
    \begin{align*}
        A ( A + (z_2-z_1) e_1^t )^{-1}
        =
        \Id
        - 
        \left( 1 + e_1^{t} A^{-1} (z_2-z_1) \right)^{-1}
        (z_2-z_1) e_1^t \cdot A^{-1}
        =
        \Id
        - 
        \left( e_1^{t} A^{-1} z_2 \right)^{-1}
        (z_2-z_1) e_1^t \cdot A^{-1}
        .
    \end{align*}
    Putting in a multiple of $\alpha z_1$, we find 
    \begin{align*}
        \alpha z_1
        -
        \left( e_1^{t} A^{-1} z_2 \right)^{-1}
        (z_2-z_1) 
        .
    \end{align*}

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior:contraction}
    Let $\calT$ be a shellable $n$-dimensional triangulation, and let the domain $\Omega \subseteq \bbR^{n}$ be the interior of the underlying set of $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling of $\calT$.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$, where $1 \leq p \leq \infty$, 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$ 
    such that 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        \leq 
        \left( 
        \Cacht{n}{k}(\calT)
        \Cachtprime{n}{k}(\calT)^{k} 
        \Cdetsieb{n}{k}(\calT)^{\frac 1 p} 
        \Csieb{n}{k}(\calT)
        \Csiebprime{n}{k}(\calT)^{k-1} 
        \Cdetacht{n}{k}(\calT)^{\frac 1 p} 
        \right)^{M}
        C_{{\PF},k,T_{0},p}
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
\end{theorem}
\begin{proof}
    We prove this via induction. 
    The statement is clearly satisfied when $M=0$ because then we only consider a single simplex,
    so that the Poincar\'e--Friedrichs inequality over a simplex can be used. 
    Let us assume that the statement is true for some integer $M-1 \geq 0$.

@@

>>>>

\begin{proof}
    The diameter of $F$ is at least as large as the height $h_S$ of some other vertex of $F$ in $T_{1}$. 
    Now, 
    \begin{align*}
%        \diam(T) \frac{ n! \vol(T) }{ \diam(T)^{n} } \leq h_S \leq \diam(F).
        \diam(T_{1}) \aspectratio(T_{1})^{-1} \leq \height(T_{1}) \leq h_S \leq \diam(F).
    \end{align*}
    The first estimate follows. 
    As for the second estimate,
    let $h_1$ and $h_2$ be the heights of $F$ in the simplices $T_1$ and $T_2$, respectively. 
    By the volume formula for simplices, $\vol(T_1) = h_1 \vol(F)/n$ and $\vol(T_2) = h_2 \vol(F)/n$, and thus $\vol(T_1)/\vol(T_2) = h_1/h_2$. Thus follows the second estimate: 
    \begin{align*}
        \frac{ \vol(T_1) }{ \vol(T_2) }
        = 
        \frac{ h_1 }{ h_2 }
        \leq
        \frac{ \diam(T_1) }{ h_2 }
        \leq
        \aspectratio(T_{1}) \frac{ \diam(F) }{ h_2 }
        \leq
        \aspectratio(T_{1}) \aspectratio(T_{2})
        .
    \end{align*}
    Lastly, let $\Xi : T_1 \rightarrow T_2$ be as stated. 
    We estimate the Lipschitz constant of $\Xi$ as follows. 
    Without loss of generality, $F$ lies in the span of the first $n-1$ coordinates and contains the origin, 
    and the unit height vector $h_0$ goes in the direction of the vertex $z_1$ is in the $n$-th coordinate direction.
    Suppose that $b, b' \in F$ and $\alpha, \alpha' \in \bbR$ such that $x = b + \alpha h_0$ and $x' = b' + \alpha' h_0$.
    Since $\Xi(h_0) = z_2$, it follows with the arguments as above and H\"older's inequality that 
    \begin{align*}
        \| \Xi(x) - \Xi(x') \|
        \leq  
        \| b - b' \| + | \alpha - \alpha' | \cdot \| z_2 \|
        \leq  
        \| b - b' \| + | \alpha - \alpha' | \aspectratio(T_1) \frac{\diam(T_2)}{\diam(T_1)} \| h_0 \|
        \leq 
        \left( 1 + \aspectratio(T_1)^2 \frac{\diam(T_2)^2}{\diam(T_1)^2} \right)^{\frac 1 2} \| x-x' \|
        .
    \end{align*}
    \color{red}
    Writing $h_0$ for the unit normal in direction $z_1$, we have 
    \begin{align*}
        \Xi(x) 
        = 
        x 
        - \frac{ \langle h_0, x \rangle }{ \langle h_0, z_1 \rangle } z_1
        + \frac{ \langle h_0, x \rangle }{ \langle h_0, z_1 \rangle } z_2
        ,
        \qquad 
        x \in \bbR^{n}
        .
    \end{align*}
    Its Jacobian is the rank-one perturbation of the linear projection along $z_1$ onto the hyperplane spanned by $F$. 
    By the Xu-Zikanatov estimate,
    \begin{align*}
        \matnorm{ \Jacobian \Xi }
        \leq 
        \frac{ \diam(T_1) }{ \height(T_1) }
        +
        \frac{ \diam(T_2) }{ \height(T_1) }
        \leq 
        \left( 1 + \aspectratio(T_2) \right)
        \aspectratio(T_1)
        .
    \end{align*}
    The desired estimates are shown.
\begin{comment}
    \mwl{Tighter estimate via geometry}
    Matrix inverse lemma:
    \begin{align*}
        A ( A + (z_2-z_1) e_1^t )^{-1}
        =
        \Id
        - 
        \left( 1 + e_1^{t} A^{-1} (z_2-z_1) \right)^{-1}
        (z_2-z_1) e_1^t \cdot A^{-1}
        =
        \Id
        - 
        \left( e_1^{t} A^{-1} z_2 \right)^{-1}
        (z_2-z_1) e_1^t \cdot A^{-1}
        .
    \end{align*}
    Putting in a multiple of $\alpha z_1$, we find 
    \begin{align*}
        \alpha z_1
        -
        \left( e_1^{t} A^{-1} z_2 \right)^{-1}
        (z_2-z_1) 
        .
    \end{align*}

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior:contraction}
    Let $\calT$ be a shellable $n$-dimensional triangulation, and let the domain $\Omega \subseteq \bbR^{n}$ be the interior of the underlying set of $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling of $\calT$.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$, where $1 \leq p \leq \infty$, 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$ 
    such that 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        \leq 
        \left( 
        \Cacht{n}{k}(\calT)
        \Cachtprime{n}{k}(\calT)^{k} 
        \Cdetsieb{n}{k}(\calT)^{\frac 1 p} 
        \Csieb{n}{k}(\calT)
        \Csiebprime{n}{k}(\calT)^{k-1} 
        \Cdetacht{n}{k}(\calT)^{\frac 1 p} 
        \right)^{M}
        C_{{\PF},k,T_{0},p}
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
\end{theorem}
\begin{proof}\color{blue}
    We first fix some notation. 
    Recall that for any $1 \leq m \leq M$, 
    the sequence $T_0, T_1, \dots, T_{m-1}$ is the shelling of a triangulation of an $n$-dimensional submanifold with boundary, 
    written $X_{m}$, whose interior domain we denote by $\Omega_m$. 
    By Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ to the triangulation completes a star in $\calT$ around some simplex $S_{m} \in X_{m-1}$, 
    and we let $\Makoto_{m-1} \subseteq X_{m-1}$ be the complement of $T_{m}$ in that newly completed star.

    We introduce the deformation mapping $\Xi_{2,m} : \patch_{\calT}(S_m) \rightarrow \Makoto_{m-1}$ of Proposition~\ref{proposition:starreflection}. 
    This mapping and its inverse are piecewise linear, and it acts as the identity along $\partial \Makoto_{m-1} \setminus \partial T$.
    We extend it to a bi-Lipschitz piecewise linear mapping $\Xi_{2} : X_{m} \rightarrow X_{m-1}$ 
    by letting it be the identity over $X_{m-1} \setminus A_{m-1}$.

    For each $T \in \calT$ of dimension $n$, there exists a sequence $1 \leq m_{T,1},\dots,m_{T,K} \leq M$, for some $1 \leq K \leq M$,
    such over the simplex $T$ we have the identity 
    \begin{align*}
        ...
    \end{align*}
\end{proof}
\begin{proof}
    We prove this via induction. 
    The statement is clearly satisfied when $M=0$ because then we only consider a single simplex,
    so that the Poincar\'e--Friedrichs inequality over a simplex can be used. 
    Let us assume that the statement is true for some integer $M-1 \geq 0$.

@@

\begin{remark}\label{remark:fullrecursivesum:exterior}
    \color{red}
    We provide a similar discussion as in Remark~\ref{remark:fullrecursivesum}, 
    this time geared towards the application of Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
    Recall that the potential $w$ for the gradient $\cartan u$ satisfies the recursive estimate 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \| \cartan u      \|_{L^{p}(T_{m})} 
        \\&\qquad
        +
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \Cfive{n}{k}(\calT)
        \Cfiveprime{n}{k}(\calT)^{k} 
        \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| \cartan u \|_{L^{p}(\Makoto_{m-1})}
        \\&\qquad
        + 
        \Cfive{n}{k}(\calT) 
        \Cfiveprime{n}{k}(\calT)^{k-1} 
        \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(\Makoto_{m-1})}
        .
    \end{align*}
    Unwrapping the recursion for the norm over the $m$-th simplex, 
    we find an estimate of the form 
    \begin{align}\label{math:fullrecursivesum:exterior}
        \| w \|_{L^{p}(T_{m})} 
        &
        \leq 
        \sum_{\ell=0}^{M} C_{m,\ell} \| \cartan u \|_{L^{p}(T_{\ell})}
        .
    \end{align}
    Here, $C_{m,\ell}$ is the coefficient of $\| \cartan u \|_{L^{p}(T_{\ell})}$, possibly zero, 
    as it appears in the unwrapped recursive estimate of $\| w \|_{L^{p}(T_{m})}$. 
    The global Poincar\'e--Friedrichs inequality follows via H\"older's inequality:
    \begin{align*}
        \| w \|_{L^{p}(\Omega)}^{p}
        &
        \leq 
        \sum_{m=0}^{M}
        \| w \|_{L^{p}(T_{m})}^{p}
        % \\&
        \leq 
        \sum_{m=0}^{M}
        \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
        \sum_{\ell=0}^{M} 
        \| \cartan u \|_{L^{p}(T_{\ell})}^{p} 
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell=0}^{M} C_{m,\ell}^{q} \right)^{\frac p q}
        \right)
        \| \cartan u \|_{L^{p}(\Omega)}^{p} 
        ,
    \end{align*}
    where $q \in [1,\infty]$ satisfies $1 = 1/p + 1/q$ and with obvious modifications if $p=1$ or $p=\infty$. 
    Alternatively, when $L$ denotes the longest path, then a different application of H\"older's inequality shows:
    \begin{align*}
        \| w \|_{L^{p}(\Omega)}^{p}
        &
        \leq 
        \sum_{m=0}^{M}
        \| w \|_{L^{p}(T_{m})}^{p}
        \\&
        \leq 
        \sum_{m=0}^{M}
        L^{\frac p q}
        \sum_{\ell=0}^{M} C_{m,\ell}^{p} \| \cartan u \|_{L^{p}(T_{\ell})}^{p} 
        \leq 
        L^{\frac p q}
        \sum_{m=0}^{M} 
        \max_{ 0 \leq \ell \leq M }
        C_{m,\ell}^{p} 
        \cdot 
        \| \cartan u \|_{L^{p}(\Omega)}^{p} 
        ,
    \end{align*}

    Hence 
    \begin{align}
        \| w \|_{L^{p}(\Omega)}^{p}
        &
        \leq 
        \sum_{m=0}^{M}
        \| w \|_{L^{p}(T_{m})}^{p}
        \\&
        \leq 
        \sum_{m=0}^{M}
        \sum_{\ell=0}^{M} 
        \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
        \delta_{m,\ell} \| \cartan u \|_{L^{p}(T_{\ell})}^{p} 
        \\&
        \leq 
        \max_{0 \leq \ell \leq M} \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
            \delta_{m,\ell} 
        \right)
        \cdot 
        \| \cartan u \|_{L^{p}(\Omega)}^{p} 
        \\&
        \leq 
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
        \right)
        \cdot 
        \| \cartan u \|_{L^{p}(\Omega)}^{p} 
        .
    \end{align}
    Now, if $q \leq p$, that is, $2 \leq p$, then 
    \begin{align}
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
        \right)^{\frac 1 p}
        \leq 
        \left(
            \sum_{\ell'=0}^{M}
            \left( \sum_{m=0}^{M} C_{m,\ell'}^{p} \right)^{\frac q p}
        \right)^{\frac 1 q}
        .
    \end{align}
    Instead if $p \leq q$, that is, $p \leq 2$, then 
    \begin{align}
        \left(
            \sum_{m=0}^{M}
            \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac p q}
        \right)^{\frac 1 p}
        \leq 
        \sum_{m=0}^{M}
        \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{q} \right)^{\frac 1 q}
        \leq 
        \sum_{m=0}^{M}
        \left( \sum_{\ell'=0}^{M} C_{m,\ell'}^{p} \right)^{\frac 1 p}
        .
    \end{align}
    So in any case, it is better than using the following Minkowski's inequality: 
    \begin{align}
        \| w \|_{L^{p}(\Omega)}
        &
        \leq 
        \left( \sum_{m=0}^{M} \left( \sum_{\ell=0}^{M} C_{m,\ell} \| \cartan u \|_{L^{p}(T_{\ell})} \right)^{p} \right)^{\frac 1 p}
        \\&
        \leq 
        \sum_{\ell=0}^{M}
        \left( \sum_{m=0}^{M} C_{m,\ell}^{p} \| \cartan u \|_{L^{p}(T_{\ell})}^{p} \right)^{\frac 1 p}
        \leq 
        \sum_{\ell=0}^{M}
        \left( \sum_{m=0}^{M} C_{m,\ell}^{p} \right)^{\frac 1 p}
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align}
\end{remark}
